Monte Carlo Linear Solvers

### 2016-01-25 doc/NCState_seminar_2014/ncstate_seminar.tex
adding nc state seminar tex file

%%---------------------------------------------------------------------------%%
%% SIAM CSE 2015 Presentation
%%---------------------------------------------------------------------------%%

@@

\title[Parallel Monte Carlo Solvers]{Parallel Algorithms for Monte Carlo
  Linear Solvers}
\author[Slattery, Hamilton, Evans]{Stuart Slattery
  \\ Steven Hamilton \\ Tom Evans}
\date[3/17/15]{March 17, 2015}
\institute[ORNL]{\small Oak Ridge National Laboratory}
\titlegraphic{\includegraphics[width=1.5in]{ORNL_Balls.pdf}}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Motivation}
  \vfill
\begin{itemize}
  \item As we move towards exascale computing, the rate of errors is expected
  to increase dramatically
  \vfill
  \begin{itemize}
  \item The probability that a compute node will fail during the course
    of a large scale calculation may be near 1
  \end{itemize}
  \vfill
\item Algorithms need to not only have increased concurrency/scalability
  but have the ability to recover from hardware faults
  \begin{itemize}
  \item Lightweight machines
  \item Heterogeneous machines
  \item Both characterized by low power and high concurrency
  \end{itemize}
  \vfill
\end{itemize}
\vfill
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Towards Exascale Concurrency and Resiliency}
  \begin{itemize}
    \item Two basic strategies:
      \vfill
      \begin{enumerate}
        \item State with current ``state of the art'' methods and make
          incremental modifications to improve scalability and fault
          tolerance
          \begin{itemize}
            \item Many efforts are heading in this direction, attempting
              to find additional concurrency to exploit
          \end{itemize}
        \vfill
        \item Start with methods having natural scalability and resiliency
          aspects and work at improving performance (e.g. Monte Carlo)
          \begin{itemize}
            \item Soft failures a component of the tally variance
            \item Hard failures potentially mitigated by replication
            \item Concurrency enabled by several levels of parallelism
          \end{itemize}
      \end{enumerate}
  \end{itemize}
  \vfill
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Outline}

  \begin{itemize}
  \item Monte Carlo Linear Solvers
    \vfill
  \item Domain Decomposition and Replication
    \vfill
  \item Scaling Studies
    \vfill
  \item Matrix-Free and Stochastic Approximate Inverse Algorithms
  \end{itemize}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Monte Carlo Synthetic Acceleration}
  \begin{itemize}
  \item Devised by Evans and Mosher in the 2000's as an acceleration
    scheme for radiation diffusion problems (LANL)
    \vfill
  \item Can be abstracted as a general linear solver with Monte Carlo as a
    preconditioner, $\mathbf{M_{MC}}$
    \vfill
  \item Combine with Richardson iteration as a ``smoother'' in between
    Monte Carlo steps:
    \begin{align*}
      \mathbf{r}^k &= \mathbf{b} - \mathbf{Ax}^k \\
      \mathbf{x}^{k+1/2} &= \mathbf{x}^k + \mathbf{r}^k \\
      \mathbf{r}^{k+1/2} &= \mathbf{b} - \mathbf{Ax}^{k+1/2} \\
      \mathbf{\delta} &= \mathbf{M_{MC}} \mathbf{r}^{k+1/2} \\
      \mathbf{x}^{k+1} &= \mathbf{x}^{k+1/2} + \delta
    \end{align*}
  \end{itemize}
\end{frame}

@@

  \center Domain Decomposition and Replication

@@

  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item Each parallel process owns a piece of the domain (linear
        system)
        \bigskip
      \item Random walks must be transported between adjacent domains
        through parallel communication
        \bigskip
      \item Domain decomposition determined by the input system
        \bigskip
      \item Load balancing not yet addressed
      \end{itemize}
    \end{column}

@@

    \begin{column}{0.5\textwidth}
      \vspace{-0.1in}
      \small
      \begin{itemize}
      \item General extension of the Milagro algorithm (LANL)
      \item Asynchronous nearest neighbor communication of histories
      \item System-tunable communciation parameters of buffer size and check
        frequency (performance impact)
      \item Need an asynchronous strategy for exiting the transport loop
        without a collective (running sum)
      \end{itemize}

      \vspace{-0.1in}
      \begin{figure}[htpb!]
        \begin{center}
          \scalebox{0.45}{ \input{domain_to_domain.pdftex_t} }
        \end{center}
      \end{figure}

@@

      \vspace{-0.3in}
      \begin{figure}[htpb!]
        \begin{center}
          \scalebox{0.85}{ \input{transport_loop.pdftex_t} }
        \end{center}
      \end{figure}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Stopping the Asynchronous Kernel}

  \vspace{-0.25in}

  \begin{figure}[htpb!]
    \begin{center}
      \scalebox{0.4}{ \input{master_comm_tree.pdftex_t} }
    \end{center}
    \caption{Master/Slave}
  \end{figure}

    \begin{column}{0.5\textwidth}
      \vspace{-0.75in}
      \begin{figure}[htpb!]
        \begin{center}
          \scalebox{0.4}{ \input{binary_comm_tree.pdftex_t} }
        \end{center}
        \caption{Binary Tree}
      \end{figure}

      \vspace{-0.5in}

      \begin{figure}[t!]
        \begin{center}
          \includegraphics[width=0.99\textwidth]{titan_weak_bvsm.pdf}
        \end{center}
        \caption{Weak scaling absolute efficiency}
        \label{fig:titan_weak_bvsm}
      \end{figure}

%%---------------------------------------------------------------------------%%
\begin{frame}{Replication}

  Different batches of Monte Carlo samples can be combined in summation via
  superposition if they have different random number streams. For two
  different batches:

  \[
  \mathbf{M_{MC}} \mathbf{x} =
  \frac{1}{2}(\mathbf{M_1}+\mathbf{M_2})\mathbf{x} 
  \]

  Consider each of these batches independent \textit{subsets} of a Monte Carlo
  operator where now the operator can be formed as a general additive
  decomposition of $N_S$ subsets:

  \[
  \mathbf{M_{MC}} = \frac{1}{N_S}\sum_{n=1}^{N_s}{\mathbf{M_n}}
  \]

  We replicate the linear problem and form each subset on a different group of
  parallel processes. Applying the subsets to a vector requires an AllReduce
  to form the sum. Each subset is domain decomposed.

  \center Scaling Studies

%%---------------------------------------------------------------------------%%
\begin{frame}{Parallel Test Application -- Nuclear Reactor Analysis}
  \small
  The simplified $P_N$ ($SP_N$) equations are an approximation
  to the Boltzmann neutron transport equation used to simulate nuclear
  reactors

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{$SP_N$ Assembly Problem}
\small
  \begin{columns}
    \begin{column}{0.4\textwidth}
      Test problem -- $3 \times 3$ array of fuel assemblies with
      control rod in center location (Profugus)
      \begin{itemize}
      \item 23 energy groups, 2 angular moments, 25M degrees of freedom
      \item 1,000 computational cores via domain decomposition
      \item We are interested in solving generalized eigenvalue problem,
        for this study we use Arnoldi as the eigensolver and compare
        different methods for solving linear systems
      \end{itemize}
    \end{column}
    \begin{column}{0.6\textwidth}
      \vspace*{-0.5in}
      \begin{figure}
        \centering
        \includegraphics[width=3in]{prob4}
      \end{figure}
    \end{column}
  \end{columns}

@@

  \center Matrix-Free and Stochastic Approximate Inverse Algorithms

@@

>>>>

%%---------------------------------------------------------------------------%%
%% Shift Roadmap Presentation
%%---------------------------------------------------------------------------%%

@@

\title[Monte Carlo Linear Solvers]{Monte Carlo Linear Solvers}
\author[Slattery, Evans, Hamilton]{Stuart Slattery \\ Tom Evans
  \\ Steven Hamilton}
\date[11/4/14]{November 4, 2014}
\institute[ORNL]{\small Oak Ridge National Laboratory}
\titlegraphic{\includegraphics[width=1.5in]{ORNL_Balls.pdf}}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Motivation}
  \vfill
\begin{itemize}
  \item As we move towards exascale computing, the rate of errors is expected
  to increase dramatically
  \vfil
  \begin{itemize}
  \item The probability that a compute node will fail during the course
    of a large scale calculation may be near 1
  \end{itemize}
  \vfill
\item Algorithms need to not only have increased concurrency/scalability
  but have the ability to recover from hardware faults
  \begin{itemize}
  \item Lightweight machines
  \item Heterogeneous machines
  \item Both characterized by low power and high concurrency
  \end{itemize}
  \vfill
\end{itemize}
\vfill
\end{frame}
%%---------------------------------------------------------------------------%%
\begin{frame}{Towards Exascale Concurrency and Resiliency}
  \begin{itemize}
    \item Two basic strategies:
      \vfill
      \begin{enumerate}
        \item State with current ``state of the art'' methods and make
          incremental modifications to improve scalability and fault
          tolerance
          \begin{itemize}
            \item Many efforts are heading in this direction, attempting
              to find additional concurrency to exploit
          \end{itemize}
        \vfill
        \item Start with methods having natural scalability and resiliency
          aspects and work at improving performance (e.g. Monte Carlo)
          \begin{itemize}
            \item Soft failures buried within the tally variance
            \item Hard failures mitigated by replication
            \item Concurrency enabled by several levels of parallelism
          \end{itemize}
      \end{enumerate}
  \end{itemize}
  \vfill
\end{frame}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Adjoint Monte Carlo}
\begin{itemize}
  \item Choose $\mathbf{P}$ and $\mathbf{W}$ such that
    $\mathbf{H}^{T} = \mathbf{P} \circ \mathbf{W}$
  \item Typical choice (Monte Carlo Almost-Optimal):
    \begin{equation*}
      \mathbf{P}_{ij} = \frac{| \mathbf{H}_{ji}| }
      {\sum_{i=1}^{N} | \mathbf{H}_{ji} |}
    \end{equation*}
  \item To estimate solution:
    \begin{itemize}
      \item Start a history in random state $i$ by sampling from distribution
        determined by $\mathbf{b}$
      \item Transition to new state $j$ based probabilities determined by
        $\mathbf{P}_i$
      \item Modify history weight based on corresponding entry in
        $\mathbf{W}_{ij}$
      \item Add contribution to $\mathbf{x}_j$ based on current history weight
    \end{itemize}
  \item A given random walk can contribute to many different components of
    the solution vector
\end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Solving the Heat Equation: Adjoint Method}

  \begin{figure}[h!]
    \begin{center}
      \includegraphics<1>[width=4in]{adjoint_1.png}
      \includegraphics<2>[width=4in]{adjoint_10.png}
      \includegraphics<3>[width=4in]{adjoint_100.png}
      \includegraphics<4>[width=4in]{adjoint_1000.png}
      \includegraphics<5>[width=4in]{adjoint_10000.png}
      \includegraphics<6>[width=4in]{adjoint_100000.png}
      \includegraphics<7>[width=4in]{adjoint_1000000.png}
      \includegraphics<8>[width=4in]{adjoint_10000000.png}
    \end{center}
    \caption{
      \only<1>{\textbf{Adjoint solution.}
        \textit{\sn{1}{0} total histories.} }
      \only<2>{\textbf{Adjoint solution.}
        \textit{\sn{1}{1} total histories.} }
      \only<3>{\textbf{Adjoint solution.}
        \textit{\sn{1}{2} total histories.} }
      \only<4>{\textbf{Adjoint solution.}
        \textit{\sn{1}{3} total histories.} }
      \only<5>{\textbf{Adjoint solution.}
        \textit{\sn{1}{4} total histories.} }
      \only<6>{\textbf{Adjoint solution.}
        \textit{\sn{1}{5} total histories.} }
      \only<7>{\textbf{Adjoint solution.}
        \textit{\sn{1}{6} total histories.} }
      \only<8>{\textbf{Adjoint solution.}
        \textit{\sn{1}{7} total histories.} }
    }
  \end{figure}
\end{frame}
%%---------------------------------------------------------------------------%%
\begin{frame}{Limitations of Monte Carlo}
\begin{itemize}
  \item Solving linear systems of equations with ``pure'' Monte Carlo methods
    is generally intractable
    \begin{itemize}
      \item Central limit theorem is barrier to accurate solutions
    \end{itemize}
\end{itemize}
\begin{figure}
  \centering
  \includegraphics[width=3.5in]{Ifiss_ConvDiff}
\end{figure}
\end{frame}

@@

  \center Residual Monte Carlo Methods

%%---------------------------------------------------------------------------%%
\begin{frame}{Sequential Monte Carlo}
  \begin{itemize}
  \item Devised by John Halton in the 1960's as a residual method
\vfill
    \item Instead of directly solving $\mathbf{Ax}=\mathbf{b}$ with Monte
      Carlo, apply Monte Carlo to residual equation
      $\mathbf{A}\delta = \mathbf{r}$
    \vfill
    \item Preconditioned Richardson iteration using Monte Carlo as
      preconditioner:
      \begin{align*}
        \mathbf{r}^{k} &= \mathbf{b} - \mathbf{Ax}^{k} \\
        \hat{\mathbf{A}} \mathbf{\delta} &= \mathbf{r}^{k} \\
        \mathbf{x}^{k+1} &= \mathbf{x}^k + \delta
      \end{align*}
    \vfill
    \item Exponential convergence achieved by decoupling Monte Carlo
      error from solution error
  \end{itemize}
\end{frame}
%%---------------------------------------------------------------------------%%
\begin{frame}{Monte Carlo Synthetic Acceleration}
  \begin{itemize}
  \item Devised by Evans and Mosher in the 2000's as an acceleration
    scheme for radiation diffusion problems (LANL)
    \vfill
  \item Can be abstracted as a general linear solver
    \vfill
  \item Combine with Richardson iteration as a ``smoother'' in between
    Monte Carlo steps:
    \begin{align*}
      \mathbf{r}^k &= \mathbf{b} - \mathbf{Ax}^k \\
      \mathbf{x}^{k+1/2} &= \mathbf{x}^k + \mathbf{r}^k \\
      \mathbf{r}^{k+1/2} &= \mathbf{b} - \mathbf{Ax}^{k+1/2} \\
      \hat{\mathbf{A}} \mathbf{\delta} &= \mathbf{r}^{k+1/2} \\
      \mathbf{x}^{k+1} &= \mathbf{x}^{k+1/2} + \delta
    \end{align*}
  \end{itemize}
\end{frame}
%%---------------------------------------------------------------------------%%
\begin{frame}{Preconditioned MCSA}
\vspace*{-0.2in}
  {
    \small
    \[
    \ve{r}^{k} = \ve{M}_L^{-1}(\ve{b}-\ve{A}\ve{M}_R^{-1}\ve{x}^{k})
    \]
    \[
    \ve{x}^{k+1/2} = \ve{x}^k + \ve{r}^k
    \]
    \[
    \ve{r}^{k+1/2} = \ve{M}_L^{-1}(\ve{b}-\ve{A}\ve{M}_R^{-1}\ve{x}^{k+1/2})
    \]
    \[
    \ve{M}_L^{-1}\ve{A}\ve{M}_R^{-1}\delta\ve{x}^{k+1/2} = \ve{r}^{k+1/2}
    \]
    \[
    \ve{x}^{k+1} = \ve{x}^{k+1/2} + \delta \ve{x}^{k+1/2}
    \]
  }
  \vspace*{-0.1in}
  \begin{itemize}
    \item Requires explicit construction of preconditioned matrix
      \begin{itemize}
        \item This has generally limited preconditioner selection to diagonal,
          block diagonal, sparse approximate inverse approaches
      \end{itemize}
    \vfill
    \item Sparsity in $\mathbf{A}$, $\mathbf{M}$, does not imply sparsity
      in $\mathbf{M^{-1}A}$
    \vfill
    \item Need to investigate preconditioning strategies that lead to sparse
      preconditioned systems (Michele Benzi, Emory)
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Residual Algorithm Comparison}
  \begin{figure}
    \includegraphics[width=3.5in]{seq_conv_100}
  \end{figure}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{3D Equilibrium Thermal Radiation Diffusion}
\vspace*{-0.1in}

{ \tiny
  T. Evans, S. Mosher, S. Slattery, S. Hamilton, ``A Monte Carlo
  synthetic-acceleration method for solving the thermal radiation
  diffusion equation,'' Journal of Computational Physics \textbf{258},
  pp.~338--358 (2014).  
}
\vspace*{-0.1in}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{figure}[ht!]
        \centerline{ \includegraphics[width=2in,clip]{mesh.pdf}}
      \end{figure}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{figure}[ht!]
        \centerline{ \includegraphics[width=2in,clip]{T_multi_mat.pdf}}
      \end{figure}
    \end{column}
  \end{columns}

\vspace*{-0.15in}
{\small
  \begin{table}
    \centering
    \begin{tabular}{cccc}
      \toprule
      \multirow{2}{*}{\bfseries Solver} &
      \bfseries Total &
      \bfseries Relative Time &
      \bfseries Relative Total \\
      & \bfseries Iterations &
      \bfseries Per Iteration &
      \bfseries Time \\
      \midrule
      CG (Jacobi) & 42,881 &  2.93 & 1.02 \\
      CG (ML) & 14,097 & 26.60 & 1.94 \\
      MCSA (Jacobi) & 90,129 & 1.0 & 1.0 \\
      \bottomrule
    \end{tabular}
  \end{table}
}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Path to Harder Problems}
  \begin{itemize}
    \item Experience with radiation diffusion suggests that for problems
      where spectral radius is low ($\lesssim 0.8$),
      MCSA can be competitive with leading methods
      \begin{itemize}
        \item Very low cost per iteration to approximate Neumann series
      \end{itemize}
    \vfill
    \item What about more challenging problems?
      \begin{itemize}
        \item Neumann series converges more slowly, so histories last longer
        \item More histories may be required
        \item What if Neumann series is not convergent?
      \end{itemize}
  \end{itemize}
\end{frame}

  \center Polynomial Formulation

%%---------------------------------------------------------------------------%%
\begin{frame}{Stopping Criteria -- Weight Cutoff}
  \begin{itemize}
    \item Histories are usually terminated when the relative weight drops
      below a specified cutoff (i.e. $\frac{w}{w_0} < \tau$)
    \vfill
    \item Adjoint MCSA on JPWH\_991 matrix with diagonal preconditioning
      ($\rho ( \mathbf{I-D^{-1}A} ) \approx 0.98$):
  \end{itemize}
\begin{figure}
  \centering
  \includegraphics[width=2.5in]{WtCutoff}
\end{figure}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Weight vs. History Length Cutoff}
\begin{figure}
  \centering
  \includegraphics[width=3in]{WtVsLength}
\end{figure}
\begin{itemize}
  \item Unlike weight cutoff, history length cutoff ``saturates'' at
    some number of histories -- fixed length Neumann series has been
    effectively reproduced
  \item Saturation point is higher for longer history lengths
\end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Monte Carlo as Approximate Polynomial Preconditioning}
  \begin{itemize}
    \item Using history length cutoff, Monte Carlo process is approximating
      a fixed length Neumann series polynomial
      \[
      x_i = \sum_{k=0}^{\infty}\sum_{i_1}^{N}\sum_{i_2}^{N}\ldots
      \sum_{i_k}^{N}h_{i,i_1}h_{i_1,i_2}\ldots h_{i_{k-1},i_k}b_{i_k}
      \]
    \vfill
    \item As number of histories grows, iteration count becomes identical to
      using ``true'' polynomial
    \vfill
    \item Why limit ourselves to the Neumann series polynomial?
      \begin{itemize}
        \item Chebyshev or GMRES polynomials are viable alternatives
      \end{itemize}
  \end{itemize}
\end{frame}
%%---------------------------------------------------------------------------%%
\begin{frame}{Neumann Series Polynomial}
\begin{table}
\caption{Adjoint MCSA with Neumann Polynomial, $1000 \times 1000$ Shifted Laplacian Matrix.
Values are MCSA iteration counts (timing in milliseconds)
\label{tab:lap_adjoint_neumann}}
\centering
\begin{tabular}{cccc}
\toprule
Histories per & \multicolumn{3}{c}{Polynomial Order} \\
\cmidrule(lr){2-4}
Iteration & 2 & 4 & 6 \\
\midrule
250 &  775(100) & 651(96) & 664(110) \\
500 &  725(122) & 502(104) & 394(95) \\
1000 & 707(174) & 482(179) & 366(144) \\
2000 & 703(280) & 471(259) & 356(251) \\
4000 & 698(494) & 467(497) & 350(458) \\
8000 & 697(923) & 464(905) & 350(873) \\
16000 & 695(1796) & 462(1768) & 347(1711) \\
\bottomrule
\end{tabular}
\end{table}
\end{frame}
%%---------------------------------------------------------------------------%%
\begin{frame}{Chebyshev Polynomial}
\begin{table}
\caption{Adjoint MCSA with Chebyshev Polynomial, $1000 \times 1000$ Shifted Laplacian Matrix.
Values are MCSA iteration counts (timing in milliseconds)
\label{tab:lap_adjoint_cheby}}
\centering
\begin{tabular}{cccc}
\toprule
Histories per & \multicolumn{3}{c}{Polynomial Order} \\
\cmidrule(lr){2-4}
Iteration & 1 & 2 & 3 \\
\midrule
250 & - & - & - \\
500 & - & - & - \\
1000 & - & - & - \\
2000 & - & 328(134) & - \\
4000 & - & 296(210) & - \\
8000 & - & 288(380) & 262(423) \\
16000 & 1132(1866) & 283(721) & 175(550) \\
\bottomrule
\end{tabular}
\end{table}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Polynomial Methods -- Summary}
  \begin{itemize}
    \item Using history length cutoff rather than weight cutoff can lead to
      large reductions in iteration counts
      \vfill
    \item A good approximation to a few terms in the Neumann series
      performs better than a statistically noisy approximation to
      full series
      \vfill
    \item Significant reduction in iteration counts are possible using
      alternate polynomials, but generally outweighed by increase in number
      of histories required
      \begin{itemize}
        \item May be very beneficial from resiliency and parallel efficiency
          standpoints
        \item Will be re-evaluated in future efforts
      \end{itemize}
  \end{itemize}
\end{frame}

  \center Parallelism

@@

  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item Each parallel process owns a piece of the domain (linear
        system)
        \bigskip
      \item Random walks must be transported between adjacent domains
        through parallel communication
        \bigskip
      \item Domain decomposition determined by the input system
      \end{itemize}
    \end{column}

@@

      \begin{itemize}
      \item General extension of the Milagro algorithm (LANL)
      \item Asynchronous nearest neighbor communication of histories
      \item Binary asynchronous communication tree for completing
        transport
      \end{itemize}

      \begin{figure}[htpb!]
        \begin{center}
          \scalebox{0.45}{ \input{domain_to_domain.pdftex_t} }
        \end{center}
      \end{figure}

@@

      \begin{itemize}
      \item Extensible to problems where histories may be created
        (i.e. variance reduction)
      \end{itemize}

      \begin{figure}[htpb!]
        \begin{center}
          \scalebox{0.45}{ \input{binary_comm_tree.pdftex_t} }
        \end{center}
      \end{figure}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Parallel Application -- Nuclear Reactor Analysis}
  \small
  The simplified $P_N$ ($SP_N$) equations are an approximation
  to the Boltzmann neutron transport equation used to simulate nuclear
  reactors

@@

\end{frame}
%%---------------------------------------------------------------------------%%
\begin{frame}{$SP_N$ Assembly Problem}
\small
  \begin{columns}
    \begin{column}{0.4\textwidth}
      Test problem -- $3 \times 3$ array of fuel assemblies with
      control rod in center location (Profugus)
      \begin{itemize}
      \item 23 energy groups, 2 angular moments, 25M degrees of freedom
      \item 1,000 computational cores via domain decomposition
      \item We are interested in solving generalized eigenvalue problem,
        for this study we use Arnoldi as the eigensolver and compare
        different methods for solving linear systems
      \end{itemize}
    \end{column}
    \begin{column}{0.6\textwidth}
      \vspace*{-0.5in}
      \begin{figure}
        \centering
        \includegraphics[width=3in]{prob4}
      \end{figure}
    \end{column}
  \end{columns}

@@

@@

  \center Conclusions

%%---------------------------------------------------------------------------%%
\begin{frame}{Additional Thoughts}
  \begin{itemize}
    \item Our current implementations rely on performing fixed number of
      histories per Monte Carlo solve
      \begin{itemize}
        \item How can we dynamically select/adapt the number of histories
          that should be performed?  Use variance-based stopping criteria?
      \end{itemize}
    \vfill
    \item The ``almost optimal'' selection of the probability and weight
      matrices is arbitrary (as long as
      $\mathbf{H}/\mathbf{H}^T=\mathbf{P} \circ \mathbf{W}$)
      \begin{itemize}
        \item If $\mathbf{P}$ is taken to have a uniform distribution within
          each row, then sampling from the distribution can be done in
          constant (rather than logarithmic) time
      \end{itemize}
    \vfill
    \item Exploring Anderson acceleration of MCSA (Tim Kelley)
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Monte Carlo Linear Solvers Library (MCLS)}

  \begin{itemize}
  \item Designed to be easily incorporated with production physics
    codes
    \bigskip
  \item General asynchronous MSOD MCSA implementation
    \begin{itemize}
    \item Forward and adjoint Monte Carlo with method of expected
      values
    \item Parallel row matrix/vector interface
    \item General fixed point iteration strategy
    \item Explicit algebraic preconditioner suite
    \item Planned extensions to threaded/hybrid parallelism
    \end{itemize}
    \medskip
  \item Implemented in C++
    \bigskip
  \item Heavy use of the Trilinos scientific computing libraries
    \bigskip
  \item Open-source BSD 3-clause license
    \bigskip
  \item \textbf{\url{https://github.com/sslattery/MCLS}}
  \end{itemize}

%%---------------------------------------------------------------------------%%
\begin{frame}{Profugus Neutronics Mini-Application}

  A mini-application provides open-source kernels that effectively
  capture the algorithmic features of full applications

  \begin{columns}
    \begin{column}{0.3\textwidth}
      \vspace*{-0.4in}
      \begin{figure}
        \centering
        \includegraphics[width=1.5in]{exnihilo.png}
      \end{figure}
    \end{column}
    \begin{column}{0.7\textwidth}
      \begin{itemize}
      \item Extracts core computational kernels from the Exnihilo transport
        code suite
        \begin{itemize}
        \item Denovo: deterministic transport solvers
        \item Shift: Monte Carlo transport
        \item Insilico: reactor neutronics (cross sections, geometry,
          input/output)
        \end{itemize}
      \item Exnihilo is export-controlled, Profugus is not
      \end{itemize}
    \end{column}
  \end{columns}

  \vspace*{0.3in}
  Profugus includes:
  \begin{itemize}
  \item Fixed source and eigenvalue problems
  \item Trilinos solvers/preconditioners
  \item Generic material and cross section libraries
  \item \textbf{\url{https://github.com/ORNL-CEES/Profugus}}
  \end{itemize}
